Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training and Inference with GPUs

### 2020-12-26 NEUROC/ms.tex
增加了Motivation各个数据集的代码
回复R1Q2.
更新R1P1
回复R1Q1.
更新Reviewer2
更新了摘要和Introduction

%! TEX root = ms.tex
%! TEX program = pdflatex
\documentclass{elsarticle}
\usepackage{geometry}
\geometry{left=3cm, right=3cm, top=2.5cm, bottom=2.5cm} 
\usepackage{lineno,hyperref}
\usepackage{amsmath, amsthm}
\usepackage{relsize}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{subfig}
\usepackage{float}
\modulolinenumbers[5]
\bibliographystyle{elsarticle-num}

@@

	\title{Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs}
    \author{Zhaokang Wang, Yunpan Wang, Chunfeng Yuan, Rong Gu$^*$ \corref{correspondingauthor1}, Yihua Huang$^*$ \corref{correspondingauthor2}}
    \cortext[correspondingauthor]{Corresponding authors with equal contribution}
    \ead{\{wangzhaokang, wangyp\}@smail.nju.edu.cn, \{cfyuan, gurong, yhuang\}@nju.edu.cn}
	\address{State Key Laboratory for Novel Software Technology, \\Department of Computer Science and Technology, Nanjing University, \\Nanjing 210023, China}

	\begin{abstract}
		The graph neural network (GNN) has become a popular research area for its state-of-the-art performance in many graph analysis tasks. 
		Recently, various graph neural network libraries have emerged.They make the development of GNNs convenient, but their performance on large datasets is not satisfying. 
		In this work, we analyze the performance bottleneck in training GNN with GPUs empirically. 
		A GNN layer can be decomposed into two parts: the vertex and the edge calculation parts. 
		According to their computational complexity, we select four representative GNNs (GCN, GGNN, GAT, GaAN) for evaluation. 
		We breakdown their training time and memory usage, evaluate the effects of hyper-parameters, and assess the efficiency of the sampling techniques.
		The experimental evaluation indicates that the edge-related calculation is the performance bottleneck for most GNNs, dominating the training time and memory usage.
		Future optimization can focus on it. The sampling techniques are essential for training big graphs on GPUs, but their current implementations still have room for improvement.
	\end{abstract}

@@

\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
\bibliography{gnnref.bib}

>>>>

% !TEX program = pdflatex
\documentclass{elsarticle}
\usepackage{geometry}
\geometry{left=3cm, right=3cm, top=2.5cm, bottom=2.5cm} 
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[colorlinks=true, allcolors=red]{hyperref}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{relsize}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{subfig}
\usepackage{float}
\usepackage{times}
\modulolinenumbers[5]
\bibliographystyle{elsarticle-num}

% Vector
\newcommand{\MyVec}[1]{\boldsymbol{#1}}
% Matrix
\newcommand{\MyMat}[1]{\boldsymbol{#1}}
% Vectors with hat
\newcommand{\hvec}[1]{\hat{\boldsymbol{h}}_{#1}}
% Model parameters
\newcommand{\Param}[1]{\textcolor{blue}{#1}}

@@

	\title{Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training and Inference with GPUs}
    \author{Zhaokang Wang, Yunpan Wang, Chunfeng Yuan, Rong Gu$^*$ \corref{correspondingauthor1}, Yihua Huang$^*$ \corref{correspondingauthor2}}
    \cortext[correspondingauthor]{Corresponding authors with equal contribution}
    \ead{\{wangzhaokang, wangyp\}@smail.nju.edu.cn, \{cfyuan, gurong, yhuang\}@nju.edu.cn}
	\address{State Key Laboratory for Novel Software Technology, \\Department of Computer Science and Technology, Nanjing University, \\Nanjing 210023, China}

	\begin{abstract}
		The graph neural network (GNN) has become a popular research area for its state-of-the-art performance in many graph analysis tasks. 
		Recently, various graph neural network libraries have emerged.They make the development of GNNs convenient, but their performance on large datasets is not satisfying. 
		In this work, we analyze the performance bottleneck in training/inferring GNN with GPUs empirically. 
		A GNN layer can be decomposed into two parts: the vertex and the edge calculation parts. 
		According to their computational complexity, we select four representative GNNs (GCN, GGNN, GAT, GaAN) for evaluation. 
		We breakdown their training time and memory usage, evaluate the effects of hyper-parameters, and assess the efficiency of the sampling techniques.
		The experimental evaluation indicates that the edge-related calculation is the performance bottleneck for most GNNs, dominating the training time and memory usage.
		Future optimization can focus on it. The sampling techniques are essential for training big graphs on GPUs, but their current implementations still have room for improvement.
	\end{abstract}

@@

%\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
\bibliography{gnnref.bib}

### 2020-12-26 NEUROC/sec1-introduction.tex
更新了摘要和Introduction

To explore the essential performance bottleneck in the GNN training, we conduct a range of experimental analysis in deep in this work. 
We focus on the efficiency bottleneck of GNN training instead of accuracy. 
We model the GNNs with the message-passing framework that decomposes a GNN layer into two parts: the vertex calculation and the edge calculation.  
According to the time complexity of the two parts, we classify the typical GNNs into four quadrants(\{high, low\} complexity $\times$ \{vertex, edge\} calculation).
We choose GCN \cite{kipf2017_gcn}, GGNN \cite{li2015_ggnn}, GAT \cite{huang2018_gat} and GaAN \cite{zhang2018_gaan} as representative GNNs of the four quadrants.

We implement them with PyG and evaluate their efficiency with six real-world datasets on a GPU.
We identify the most time-consuming stage in the GNN training by decomposing the training time per epoch from the layer level to the operator level.
We also analyze the memory usage during the training to discover the main factor that limits the data scalability of GNN training on GPUs. 
Finally, we evaluate whether or not the sampling techniques affect the performance bottleneck. 
The key findings and insights are summarized below.

\begin{itemize}
    \item \textbf{The training time and the memory usage of a GNN layer is mainly affected by the dimensions of the input/output hidden feature vectors.}
    Fixing other hyper-parameters, the training time and the memory usage of a GNN layer increase linearly as the input/output dimensions.
    \item \textbf{The edge-related calculation is the performance bottleneck for most GNNs.}
    For GNNs with high edge calculation complexity, most of the training time is spent on conducting the messaging function for every edge. 
    For GNNs with low edge calculation complexity, the message collection and aggregation consumes most of the training time.
    \item \textbf{The high memory usage of the edge calculation is the main factor limiting the data scalability of GNN training.}
    The edge calculation generates and caches many intermediate results. 
    They are an order of magnitude larger than the dataset itself. 
    As GPUs have limited on-chip memory, the high memory consumption prevents us from training big graphs on GNNs.
    \item \textbf{The sampling techniques can significantly reduce training time and memory usage.}
    They are essential for training big graphs on GPUs. 
    However, the existing implementation is still inefficient. 
    Under big batch sizes, the time spent on the sampling may exceed the time spent on the training. 
    Under small batch sizes, the sampled graphs are also small, wasting the computing power of GPUs.
\end{itemize}

Based on the insights, we provide several potential optimization directions:
\begin{itemize}
      \item To reduce training time, \textbf{optimizations should focus on improving the efficiency of the edge calculation}. 
      One may consider developing optimized operators for the messaging step that is the major source of computing costs in the edge calculation.
       Fusing operators of the collection step, messaging function and the aggregation step together is another way to reduce the overheads in the edge calculation.
      \item To reduce memory usage, \textbf{optimizations should focus on reducing the intermediate results in the edge calculation}. 
      One may consider adopting the checkpoint mechanism to cache less intermediate results during the forward phase and re-calculate the needed data on the fly during the backpropagation.
      \item To improve the efficiency of the sampling techniques, \textbf{one may consider overlapping the sampling on the CPU side with the training on the GPU side.} 
      Choosing a proper batch size automatically is another potential optimization.
\end{itemize}
We hope that our analysis can help the developers of the GNN libraries/systems have a better understanding of the characteristics of GNN training and propose more targeted optimizations.

>>>>

To explore the essential performance bottleneck in the GNN training/inference, we conduct a range of experimental analysis in deep in this work. 
We focus on the efficiency bottleneck of GNN training/inference.
We model the GNNs with the message-passing framework that decomposes a GNN layer into two parts: the vertex calculation and the edge calculation.  
According to the time complexity of the two parts, we classify the typical GNNs into four quadrants(\{high, low\} complexity $\times$ \{vertex, edge\} calculation).
We choose GCN \cite{kipf2017_gcn}, GGNN \cite{li2015_ggnn}, GAT \cite{huang2018_gat} and GaAN \cite{zhang2018_gaan} as representative GNNs of the four quadrants.

We implement them with PyG and evaluate their efficiency with six real-world datasets on a GPU.
We identify the most time-consuming stage in the GNN training/inference by decomposing the training/inference time per epoch from the layer level to the operator level.
We also analyze the memory usage during the training/inference to discover the main factor that limits the data scalability of GNN training/inference on GPUs. 
Finally, we evaluate whether or not the sampling techniques affect the performance bottleneck. 
The key findings and insights are summarized below.

\begin{itemize}
    \item \textbf{The training/inference time and the memory usage of a GNN layer is mainly affected by the dimensions of the input/output hidden feature vectors.}
    Fixing other hyper-parameters, the training/inference time and the memory usage of a GNN layer increase linearly as the input/output dimensions.
    \item \textbf{The edge-related calculation is the performance bottleneck for most GNNs.}
    For GNNs with high edge calculation complexity, most of the training/inference time is spent on conducting the messaging function for every edge. 
    For GNNs with low edge calculation complexity, the message collection and aggregation consumes most of the training/inference time.
    \item \textbf{The high memory usage of the edge calculation is the main factor limiting the data scalability of GNN training/inference.}
    The edge calculation generates and caches many intermediate results. 
    They are an order of magnitude larger than the dataset itself. 
    As GPUs have limited on-chip memory, the high memory consumption prevents us from training/inferring big graphs on GNNs.
    \item \textbf{The sampling techniques can significantly reduce training/inference time and memory usage.}
    They are essential for training big graphs on GPUs. 
    However, the existing implementation is still inefficient. 
    Under big batch sizes, the time spent on the sampling may exceed the time spent on the training/inference. 
    Under small batch sizes, the sampled graphs are also small, wasting the computing power of GPUs.
\end{itemize}

Based on the insights, we provide several potential optimization directions:
\begin{itemize}
      \item To reduce training/inference time, \textbf{optimizations should focus on improving the efficiency of the edge calculation}. 
      One may consider developing optimized operators for the messaging step that is the major source of computing costs in the edge calculation.
       Fusing operators of the collection step, messaging function and the aggregation step together is another way to reduce the overheads in the edge calculation.
      \item To reduce memory usage, \textbf{optimizations should focus on reducing the intermediate results in the edge calculation}. 
      One may consider adopting the checkpoint mechanism to cache less intermediate results during the forward phase and re-calculate the needed data on the fly during the backpropagation.
      \item To improve the efficiency of the sampling techniques, \textbf{one may consider overlapping the sampling on the CPU side with the training on the GPU side.} 
      Choosing a proper batch size automatically is another potential optimization.
\end{itemize}
We hope that our analysis can help the developers of the GNN libraries/systems have a better understanding of the characteristics of GNN training/inference and propose more targeted optimizations.
